---
title: "SIF Aldabra tidal predictions"
resource_files:
- data/processed/tide_pred/loc_1/2016.rds
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "charts", align: left}
      - { title: "calendar", href: "/aldabratides_calendar", align: left }
      - { title: "download", href: "/aldabratides_download", align: left }
    orientation: columns
    css: tides_style.css
    theme: yeti
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(magrittr)
library(plotly)
library(ggplot2)
library(RSQLite)
library(DT)

tide_models_params <- readRDS("data/processed/tide_models_params.rds")
source("dashboard_functions.R")
tide <- dplyr::src_sqlite('data/processed/tide_predictions.sqlite3')

# local coordinates of Aldabra
coords <- matrix(c(46.3, -9.4), nrow = 1)
```

Options {.sidebar}
-----------------------------------------------------------------------

```{r inputs}
selectInput("Loc", label = h3("Location"), 
            choices = list("Picard, Research Station" = 1), 
            selected = 1)

dateRangeInput("DateRange", label = h4("Time Range"), start = Sys.Date(), end = Sys.Date(), min = NULL, max = NULL, format = "d MM", startview = "month", weekstart = 0, language = "en", separator = " to ")


# selectInput("TableInterval", label = h4("Table Interval"), 
#             choices = list("High/Low" = 0, "High only" = 1, "Low only" = -1, 
#                            "15 min" = 15, "30 min" = 30,
#                            "60 min" = 60), 
#             selected = 0)

```

```{r nuts and bolts}
# constants
c_t <- as.POSIXct(as.POSIXlt(Sys.time(), tz = "Indian/Mahe"))
little_extra <- 10*60 # padding around each date to calculate high and low tides

# reactive parts
datum <- reactive(tide_models_params[[as.integer(input$Loc)]]$features1[1])
msl <- reactive(tide_models_params[[as.integer(input$Loc)]]$msl - datum())

# filter water height for a time interval
read_height <- function(dr, loc){
	d_start <- as.integer(dr[1])
	d_end <- as.integer(dr[2])
	o <- dplyr::tbl(tide, 'tide_prediction') %>%
    dplyr::filter(date >= d_start,
                  date <= d_end, 
    							loc_id == loc) %>%
		dplyr::collect()
	
	o %>%
    dplyr::mutate(time = as.POSIXct(date, tz = "Indian/Mahe", origin = "1970-01-01"))
}


# baseline height
day_height <- reactive({
  dr <- input$DateRange
  l <- as.integer(input$Loc)

  dr <- sort(dr)
  dr[2] <- dr[2] + 1
  dr[1] <- dr[1]
  dr <- as.POSIXct(trunc(as.POSIXct(as.POSIXlt(as.POSIXct(dr), tz = "Indian/Mahe")), "days"))
  dr[1] <- dr[1] - little_extra
  dr[2] <- dr[2] + little_extra
  
  read_height(dr, l)
})

read_high_low <- function(dr, loc) {
	
	d_start <- as.integer(dr[1])
	d_end <- as.integer(dr[2])
	o <- dplyr::tbl(tide, 'tide_high_low') %>%
    dplyr::filter(date >= d_start,
                  date <= d_end, 
    							loc_id == loc) %>%
		dplyr::collect()
	
	o %>%
    dplyr::mutate(time = as.POSIXct(date, tz = "Indian/Mahe", origin = "1970-01-01"))
}

day_high_low <- reactive({
	dr <- input$DateRange
	l <- as.integer(input$Loc)
	
	dr <- sort(dr)
  dr[2] <- dr[2] + 1
  dr[1] <- dr[1]
  dr <- as.POSIXct(trunc(as.POSIXct(as.POSIXlt(as.POSIXct(dr), tz = "Indian/Mahe")), "days"))
  dr[1] <- dr[1] - little_extra
  dr[2] <- dr[2] + little_extra
  
  read_high_low(dr, l)
})


# table
out_table <- reactive({

    h <- day_high_low()

    h %>% 
      format_tide_table() %>%
      datatable(extensions = 'Buttons',
                options = list(dom = 'Bt',
                               buttons = c('copy', 'csv', 'pdf', 'print'),
                               bPaginate = FALSE), 
                rownames = FALSE,
                colnames = c('height [m]' = 'height')) %>%
      return()
  
})

```



Column {data-width=650}
-----------------------------------------------------------------------

###

```{r tide chart}
renderPlotly({
	pdf(NULL)
  h <- day_height() %>%
  	dplyr::filter(lubridate::minute(time) %% 5 == 0) %>%
    dplyr::mutate(time_text = format(time, "%d/%m/%y %I:%M%p", tz = "Indian/Mahe"),
                  h = paste(height, "m"))

  p <- h %>%
    dplyr::mutate(boating = cut(height, c(-10, 1.6, 2, 10))) %>%
    ggplot(aes(x = time, y = height, text = time_text))
  
  # if (difftime(max(h$time), min(h$time), units = "day") < 2) {
  #   p <- p + 
  #     geom_tile(aes(y = msl(), fill = boating), height = 0.05, alpha = 0.75) +
      # scale_fill_manual(values = c("#e41a1c", "#ccebc5", "#4daf4a")) +
  #     scale_y_continuous(expand = c(0,0)) 
  # }
  
  p <- p +
    geom_line(aes(group = 1)) +
    geom_ribbon(aes(ymax = height, ymin = msl(), group = 1), alpha = 0.5, fill = "#DAE8EC") + theme_bw() +
    # theme(panel.background = element_rect(fill = "#f2f2f2")) +
    scale_x_datetime(date_minor_breaks = "1 hour", expand= c(0,1)) +  
    # scale_colour_manual(values = c("#e41a1c", "#ccebc5", "#4daf4a")) +
    xlab("time") +ylab("height [m]") +
    geom_vline(xintercept = c_t, colour = "green") +
    theme(legend.position = "none")
  
  ggplotly(p, tooltip = c("text","y")) %>% config(displayModeBar = F)
})
```


> All times displayed in Seychelles time (SCT +0400Z) — Tidal predictions are calculated on five minutes intervals — High and low times computed from the first and second order derivatives of the predictions — Water height levels are based on a MLWS (Mean Low Water Springs) datum — Water depth was computed based on latitude and hydrostatic pressure using the UNESCO 1983 polynomial approximation for a salinity of 35 parts per thousand and a termperature of 0°C. Calculations use algorithms implemented in   [marelac](https://cran.r-project.org/web/packages/marelac/index.html) — Tidal models were selected comparing the AIC values of candidate models using the  4, 7, 37 and 60th most common constituents. The slowest harmonic included was the annual solar term, and  therefore, for each site, predictions need to be taken with care until at least once one year of data has been collected. Tidal models included daily nodal corrections and were calculated using  [TideHarmonics](https://cran.r-project.org/web/packages/TideHarmonics/index.html) — Design & Implementation by [Fernando Cagua](mailto:fernando@cagua.co)


Column {data-width=350}
-----------------------------------------------------------------------

### Current water height

```{r}

current_tide <- reactive({
	l <- as.integer(input$Loc)
  h <- read_height(c(c_t - 7*3600, c_t + 7*3600), l)
  current_h <- dplyr::slice(h, round(nrow(h)/2))
  read_high_low(c(c_t - 7*3600, c_t + 7*3600), l) %>%
    dplyr::bind_rows(current_h) %>%
    dplyr::arrange(time) %>%
    dplyr::slice(seq(which(is.na(l_m_m))-1, which(is.na(l_m_m))+1))
})

renderGauge({
  cur_ti <- current_tide()

  if(cur_ti$height[1] < cur_ti$height[3]){
  water_is <- "→ raising →"
} else {
  water_is <- "← falling ←" 
}

hl <- sort(cur_ti$height[c(1,3)])

gauge(round(cur_ti$height[2], 2), 
      min = round(hl[1], 2), 
      max = round(hl[2], 2),
      abbreviate = T, abbreviateDecimals = 2, 
      symbol = "m", 
      label = water_is,
      sectors = gaugeSectors(success = c(2, 10), warning = c(1.6, 2), danger = c(-10, 1.6), colors = c("#78B976", "#D0E7CB", "#E04E5A")))
})


tide_text <- reactive({
  cur_ti <- current_tide() %>%
    dplyr::mutate(h_l = mapvalues(h_l, c(TRUE, FALSE), c("high", "low")),
                  time = format(time, "%I:%M %p"))
  
  paste("last", cur_ti$h_l[1], "tide was at", cur_ti$time[1], "and the next",
        cur_ti$h_l[3], "tide will be at", cur_ti$time[3])
})
```

> The `r renderText(tide_text())`

###

```{r}
renderDataTable(out_table())
```
